---
title: "Homework8"
author: "Daniel Wu (EID: djw3627)"
date: "2025-04-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

To access my GitHub repository, click here: [https://github.com/DanielWu3627/SDS315](https://github.com/DanielWu3627/SDS315). Please check the file named **Homework8.Rmd**.

```{r}
library(ggplot2)
library(mosaic)
setwd("/Users/danielwu/Documents/Study/SDS 315")
cre <- read.csv("creatinine.csv")
covid <- read.csv("covid.csv")
milk <- read.csv("milk.csv")
```

# Problem 1
```{r}
ggplot(cre, aes(x=age, y=creatclear)) + 
  geom_point() + 
  geom_smooth(method='lm', color='black') + labs(title = "Age and Creatine Clearance Rate in ml/minute") 

lm0 = lm(creatclear ~ age, data=cre)

coef(lm0) %>% 
  round(2)
```
## Part A
For a 55-year-old, we expect the creatinine clearance should be about 113.71 ml/minute. I determined this by first running a regression analysis of creatinine clearance vs. age and then using the coef() function to get the y-intercept, which is about 147.81, and the slope, which is about -0.62. Therefore, with y = slope * x + y-intercept (y = -0.62x + 147.81) is -0.62 * 55 + 147.81 = 113.71 ml/minute.

## Part B
Creatinine clearance decreases by a rate of approximately -0.62 ml/minute per year. I determined this by first running a regression analysis and then using the coef() function to get the slope, which is about -0.62. As age increases by 1, the creatinine clearance should decrease by 0.62 ml/minute.

## Part C
y = -0.62x + 147.81 -> -0.62 * 40 + 147.81 = 123.01 (expected creatine value for a 40-yr-old)
y = -0.62x + 147.81 -> -0.62 * 60 + 147.81 = 110.61 (expected creatine value for a 60-yr-old)
The expected creatine clearance rate is 123.01 ml/minute for the 40-yr-old. This 40-yr-old has a rate of 135, which is about 11.99 above the expected value. The expected creatine rate for a 60-yr-old is 110.61 ml/minute, but that 60-yr-old's is 112, with a difference of 1.39 above the expected value. Therefore, the creatine clearance rate is healthier for the 40-year old with a rate of 135.


# Problem 2

## 1.

```{r}
spain <- covid[covid$country == "Spain",]
italy <- covid[covid$country == "Italy",]
ggplot(italy, aes(x=days_since_first_death, y=log(deaths))) + 
  geom_point() +
  geom_smooth(method="lm",  se=FALSE, color=2) + labs(title = "Number of reported COVID deaths in Italy")

covid_double_italy = do(10000)*lm(log(deaths) ~ days_since_first_death, data=mosaic::resample(italy))
confint(covid_double_italy, level = 0.95)
double_italy <- round((70 / ((0.180)*100)),1)
paste0("The doubling time for Italy is approximately ", double_italy)
```

The estimated growth rate for Italy is 0.180, with a 95% confidence interval [0.159, 0.208]. The estimated doubling time for Italy is 3.9 days, with a 95% confidence interval [`r round((70 / ((0.208)*100)),1)`, `r round((70 / ((0.159)*100)),1)`].

```{r}
ggplot(spain, aes(x=days_since_first_death, y=log(deaths))) + 
  geom_point() +
  geom_smooth(method="lm", se=FALSE, color=2) + labs(title = "Number of reported COVID deaths in Spain")

covid_double_spain = do(10000)*lm(log(deaths) ~ days_since_first_death, data=mosaic::resample(spain))
confint(covid_double_spain, level = 0.95)
double_spain <- round((70 / ((0.310)*100)),1)
paste0("The doubling time for Spain is ", double_spain)
```

The estimated growth rate is 0.287, with a 95% confidence interval [0.235, 0.317]. The estimated doubling time for Spain is 2.3 days, with a 95% confidence [`r round((70 / ((0.317)*100)),1)`, `r round((70 / ((0.234)*100)),1)`].




\newpage

```{r}
ggplot(covid, aes(x=days_since_first_death, y=deaths, color = country)) + 
  geom_line() + 
  labs(title = "Number of reported COVID deaths over time in Italy and Spain")
```

# Problem 3
```{r}
milk_demand = do(10000)*lm(log(sales) ~ log(price), data=mosaic::resample(milk))
confint(milk_demand, level = 0.95)
```

I bootstrapped 10,000 times. For each time, I performed regression analysis for log(sales) vs. log(price). The coefficient for log(price) is the exprected price elasticity, which is around -1.46. When the price increases by 1 %, the sales decreases at the 95% confidence level of [-1.78, -1.46].